Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-11-28 Ben_Manuscripts/structure_paper/Outline.tex
Added comments on outline.

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item At best, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes.
  \begin{itemize}
    \item Tune size and functionality of building blocks to control pore
    size and shape: One can accomplish targeted separations with high 
    selectivity by tuning shape, size and functionality of the molecular
    building blocks which form these materials. % BJC: "these materials" --> "nanostructured membranes", or is that redundant?
    \item As a result, solute rejecting pores can have their sizes tuned
    uniformly, resulting in strict size cut-offs.
    \item Entirely different mechanisms may govern transport in a given
    nanostructured material which can inspire novel separation techniques.
  \end{itemize}

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 have been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    % BJC: New paragraph?
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in low flux membranes, slowing research in the field.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated.
  \end{itemize}

  % BJC: Not sure this figure needs to be in the paper
  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{soft_confinement.png}
	\caption{The wedge shaped liquid crystal monomer (a) self assembles into 
	mesophases with hexagonally packed pores (b). The pores are made of 
	stacked monomer disks. A sub-micron-thick film is created by casting a 
	dilute solution of Na-GA3C11/THF solution onto a silicon substrate and 
	and allowing the solvent to evaporate. The thin film contains nanoporous 
	columns which lie parallel to the film plane. (d) When a soft PDMS pad is 
	imposed to the thin film, with subsequent thermal annealing, the columns
	align perpendicular to the film plane. (e) Photo-cross-linking of the
	aligned film creates a mechanically stable thin film with vertically 
	aligned nanopores}~\label{fig:soft}
  \end{figure}

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport}~\label{fig:detail}
  \end{figure}

  Our approach to constructing a general model will follow the development of a
  model of a specific LLC membrane with sufficient experimental characterization.
    \begin{itemize}
	    \item We have chosen to focus on assemblies formed by Na-GA3C11
	    \item We have also narrowed our scope to the development of 
	    a model of the Col\textsubscript{h} phase membrane.
	    \item Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
	    phase is a simpler starting point, due to the absence of water, and has
	    more detailed experimental structural data.  % BJC: More detailed because of WAXS. HII doesn't have any (that I know of) WAXS
	    \item Despite having structural data, there is still information which 
	    experiment cannot definitively answer.
    \end{itemize}

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. We want to
  know (1) If layers do exist, how many monomers constitute a single layer?
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

  (2) Does our model support the existence of layers and if so, how well defined
  are the layers?
  \begin{itemize}
       \item Experimentally, their existence is supported by evidence of strong 
       $\pi$-$\pi$ stacking interactions in the direction perpendicular to the
       membrane plane.
       \item $\pi$-$\pi$ stacking will only occur between the aromatic monomer head groups which
       leaves no description of what is happening in the monomer tail region
       \item The tails may entangle isotropically while stacking order is maintained
       among headgroups. 
  \end{itemize}

   (3) How do monomers in each layer position themselves with respect to surrounding 
   layers?
%     Even if we knew the number of monomers in each layer, we still would not know how 
%     monomers in each layer are positioned with respect to other layers. 
   \begin{itemize}
      \item The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly 
      in this system\cite{gazit_possible_2002} %is thought to be $\pi$-$\pi$ stacking interactions between aromatic headgroups \cite{gazit_possible_2002}. 
      \item Gas phase ab initio studies of benzene dimers have shown a clear energetic
      advantage for parallel displaced and T-shaped $\pi$-$\pi$ stacking conformations versus a
      sandwiched conformation ~\cite{sinnokrot_estimates_2002}.
      \item Substituted benzene rings exhibit an even stronger $\pi$-$\pi$ stacking 
      attraction which favors the parallel displaced configuration in all cases
      except where the substitutions are extremely electron withdrawing
      \cite{waller_hybrid_2006,ringer_effect_2006}.
      \item We can use simulated X-ray diffration patterns to compare the two 
      stacking configurations. 
   \end{itemize}

   (4) Is it necessary to include any water in order to appropriately model the 
   Col\textsubscript{h} phase?
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
   \end{itemize}

%     \item While we might be able to provide answers to these questions using a 
%     molecular model, there remains the possibility that there is more than one 
%     metastable state associated with a given LLC system.

@@

  \begin{itemize}
    \item We validate the model using as much experimental information as possible.
    \item We are most interested in reproducing the conclusions about structure
    which have been made from X-ray diffraction (XRD) experiments and in matching ionic
    conductivity measurements~\cite{feng_thin_2016}.
    \begin{itemize}
	    \item We have compared simulated X-ray diffraction patterns to experiment in 
	    order to match major features present in the 2D patterns.
	    \item We calculated ionic conductivity using two agreeing methods.
	    \item We examined the the influence of crosslinking on membrane structure.
    \end{itemize}
    \item The structure-building approach and analysis used in this paper can be readily extended
    to the H\textsubscript{II} phase and other similar LC systems.
  \end{itemize}

  \section*{Methods}

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using tools from Openeye Scientific Software. All molecular dynamics 
  simulations were run using the latest version of Gromacs 2016. 
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned using the am1bccsym method of molcharge
    shipped with Openeye Scientific software's QUACPAC % BJC: how to cite?
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    \item We chose initial guesses for the remaining structural parameters based on
    experimental data and treated them as variables during model development.
  \end{itemize}

  We used experimental wide angle X-ray scattering (WAXS) data (produced as described 
  in~\cite{feng_scalable_2014}) and small angle X-ray scattering (SAXS) data from~\cite{feng_thin_2016}
  to inform our choices of some initial structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily
  on the 2D WAXS data since it encodes all structural details down to the sub-nm scale.
  \begin{itemize}
        \item There are five major features of interest present in the 2D experimental
        pattern shown in Figure~\ref{fig:WAXS}.
        \item The first is located at $q_z$ = 1.7 \angstrom$^-1$,
        corresponding to a real spacing of 3.7 \angstrom~. The reflection is
        attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
        perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
        reflection will be referred to as R-$\pi$.
        \item A weak intensity line is located at exactly half the $q_z$ value of
        R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a
        real space periodic spacing of 7.4 \angstrom~. This reflection has been
        interpreted  as 2\textsubscript{1} helical ordering of aromatic rings
        along the z axis meaning if the positions of the aromatic rings can
        be traced by a helix, then for each turn in the helix, there should be
        two aromatic rings. For this reason it will be referred to as R-helix.
        \item A third major reflection is marked by a low intensity ring located
        at r = 1.4 \angstrom$^-1$. The real space separation
        corresponds to 4.5 \angstrom~ which is characteristic of the average
        spacing between packed alkane chains. This reflection will be called R-alkanes.
        \item Within R-alkanes, are four spots of higher relative intensity which
        will be called R-spots. All are located $\approx 40$ degrees from the $q_z$ axis
        in their respective quadrants. In many liquid crystal systems this can be
        explained by the tilt angle of the alkane chains with respect to the xy plane. % BJC: References
        \item The final feature corresponds to the spacing and symmetry of
        the d\textsubscript{100} plane which can be related to the distance between
        pores. The feature, which will be called R-pores, is characterized by dots
        along $q_z$ = 0. The spacing between dots is indicative of the hexagonal
        symmetry of the packed pores. The same information at higher resolution is obtained using a SAXS
        setup. By radially integrating the 2D data one gets a 1D curve which is 
        shown in Figure~\ref{fig:SAXS}.  % BJC: I also have 2D SAXS data
  \end{itemize}

@@

  We chose the initial layer spacing based on R-$\pi$.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be parallel to the xy plane.
    \item The layers are placed so aromatic rings are stacked 3.7 \angstrom~
    apart in the z-direction.
    \item To quantify the degree of layering in our system, we calculated
    a spatial correlation function, $g(z)$, measured along the z-axis 
    (perpendicular to the membrane plane)
    \item To calculate $g(z)$, we binned the z component distances between 
    the center of mass of each benzene ring and all others of the same pore 
    over 50 ns of equilibrated trajectory and then normalized by the average
    number density.
    \item We compare the degree of layering between systems based on the
    amplitude of the first peak in $g(z)$. We halve the difference between the
    maximum of the first peak and the following minimum. We compare the difference
    to the mean to get a percentage deviation from the average number density.
    \item To extract the average distance between layers we applied a discrete 
    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

>>>>

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
%MRS2: couldn't you argue that the fact that they diffuse and dissolve at different rates means that it makes it possible to  design targeted separations?
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item At best, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
%MRS2: is ``well-defined'' the right word if there is a ``pore size distribution''?
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes.
  \begin{itemize}
    \item Tune size and functionality of building blocks to control pore
    size and shape: One can accomplish targeted separations with high 
    selectivity by tuning shape, size and functionality of the molecular
    building blocks which form these materials. % BJC: "these materials" --> "nanostructured membranes", or is that redundant? %MRS2: seems fine for now, we can tweak this later. 
    \item As a result, solute rejecting pores can have their sizes tuned
    uniformly, resulting in strict size cut-offs.
    \item Entirely different mechanisms may govern transport in a given
    nanostructured material which can inspire novel separation techniques.
  \end{itemize}

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 have been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    % BJC: New paragraph?
    %MRS2: shorter paragraphs generally good.
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in low flux membranes, slowing research in the field.
%MRS2: slowed research because it wasn't clear there was a route to scaleable filtration.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated.
%MRS2: not THAT many people are doing it, so hard to say reinvigorated quite yet. Pick other verb.
  \end{itemize}

  % BJC: Not sure this figure needs to be in the paper. 
  %MRS2: good question.  It helps give people context, which is necessary to move forward.  Perhaps 
  %leave it in for now, though may need to get permission to reprint it.
  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{soft_confinement.png}
	\caption{The wedge shaped liquid crystal monomer (a) self assembles into 
	mesophases with hexagonally packed pores (b). The pores are made of 
	stacked monomer disks. A sub-micron-thick film is created by casting a 
	dilute solution of Na-GA3C11/THF solution onto a silicon substrate and 
	and allowing the solvent to evaporate. The thin film contains nanoporous 
	columns which lie parallel to the film plane. (d) When a soft PDMS pad is 
	imposed to the thin film, with subsequent thermal annealing, the columns
	align perpendicular to the film plane. (e) Photo-cross-linking of the
	aligned film creates a mechanically stable thin film with vertically 
	aligned nanopores}~\label{fig:soft}
  \end{figure}

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
%MRs: specify what type of modification.
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size 
%MRS2: controllably
      or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
%MRS2: not quite sure of the phrasing here on next two points. Something more a long the lines of the choice of head group 
may allow tuning rejection of solutes based on size?
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  %MRS2: provenance of the atomistic figure could be explained better in the caption. It also looks disordered. Will eventually need a higher resolution figure for the previous understanding pore.
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport}~\label{fig:detail}
  \end{figure}

  Our approach to constructing a general model will follow the development of a
  model of a specific LLC membrane with sufficient experimental characterization.
    \begin{itemize}
	    \item We have chosen to focus on assemblies formed by Na-GA3C11
	    \item We have also narrowed our scope to the development of 
	    a model of the Col\textsubscript{h} phase membrane.
	    \item Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
	    phase is a simpler starting point, due to the absence of water, and has
	    more detailed experimental structural data.  % BJC: More detailed because of WAXS. HII doesn't have any (that I know of) WAXS
%MRS2: explicitly state the presence of WAXS data.
	    \item Despite having structural data, there is still information which 
	    experiment cannot definitively answer.
            %MRS2: such as?
    \end{itemize}

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. We want to
  know 
\begin{enumerate}
\item If layers do exist, how many monomers constitute a single layer? \label{point:monomernum}
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

 \item Does our model support the existence of layers and if so, how well defined
  are the layers? \label{point:layers}
  \begin{itemize}
       \item Experimentally, their existence is supported by evidence of strong 
       $\pi$-$\pi$ stacking interactions in the direction perpendicular to the
       membrane plane.
       \item $\pi$-$\pi$ stacking will only occur between the aromatic monomer head groups which
       leaves no description of what is happening in the monomer tail region
       \item The tails may entangle isotropically while stacking order is maintained
       among headgroups. 
  \end{itemize}

   \item How do monomers in each layer position themselves with respect to surrounding 
   layers? \label{point:orientation}
%     Even if we knew the number of monomers in each layer, we still would not know how 
%     monomers in each layer are positioned with respect to other layers. 
   \begin{itemize}
      \item The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly 
      in this system\cite{gazit_possible_2002} %is thought to be $\pi$-$\pi$ stacking interactions between aromatic headgroups \cite{gazit_possible_2002}. 
      \item Gas phase ab initio studies of benzene dimers have shown a clear energetic
      advantage for parallel displaced and T-shaped $\pi$-$\pi$ stacking conformations versus a
      sandwiched conformation ~\cite{sinnokrot_estimates_2002}.
      \item Substituted benzene rings exhibit an even stronger $\pi$-$\pi$ stacking 
      attraction which favors the parallel displaced configuration in all cases
      except where the substitutions are extremely electron withdrawing
      \cite{waller_hybrid_2006,ringer_effect_2006}.
      \item We can use simulated X-ray diffration patterns to compare the two 
      stacking configurations. 
   \end{itemize}

   \item Is it necessary to include any water in order to appropriately model the 
   Col\textsubscript{h} phase? \label{point:water}
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
   \end{itemize}
    \end{enumerate}
%MRS2: size and how well well defined the pore is (how suddenly the RDF changes in the pore) is also a question.

%MRS2: hmm.  This last point still seems somewhat relevant.  Why left out?
%     \item While we might be able to provide answers to these questions using a 
%     molecular model, there remains the possibility that there is more than one 
%     metastable state associated with a given LLC system.

@@

  \begin{itemize}
    \item We validate the model using as much experimental information as possible.
    \item We are most interested in reproducing the conclusions about structure
    which have been made from X-ray diffraction (XRD) experiments and in matching ionic
    conductivity measurements~\cite{feng_thin_2016}.
    \begin{itemize}
	    \item We have compared simulated X-ray diffraction patterns to experiment in 
	    order to match major features present in the 2D patterns.
	    \item We calculated ionic conductivity using two agreeing methods.
	    \item We examined the the influence of crosslinking on membrane structure.
    \end{itemize}
    \item The structure-building approach and analysis used in this paper can be readily extended
    to the H\textsubscript{II} phase and other similar LC systems.
  \end{itemize}
%MRS2: and potentially used to help design membranes.  
  \section*{Methods}

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using tools from Openeye Scientific Software. 
%MRS2: be specific what tools
All molecular dynamics 
  simulations were run using the latest version of Gromacs 2016. %MRS2: give specific version, since latest may be different when this is published.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned using the am1bccsym method of molcharge
    shipped with Openeye Scientific software's QUACPAC % BJC: how to cite? %MRS2: webpage if no publication, 
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    %MRS2: I was referring to the data choice of 20 vs 10.  Yes, you can refer to it here, because the expectation is that peple will only look at it after reading the whole paper.
    \item We chose initial guesses for the remaining structural parameters based on
    experimental data and treated them as variables during model development.
  \end{itemize}

  We used experimental wide angle X-ray scattering (WAXS) data (produced as described 
  in~\cite{feng_scalable_2014}) and small angle X-ray scattering (SAXS) data from~\cite{feng_thin_2016}
  to inform our choices of some initial structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily
  on the 2D WAXS data since it encodes all structural details down to the sub-nm scale.
  \begin{itemize}
        \item There are five major features of interest present in the 2D experimental
        pattern shown in Figure~\ref{fig:WAXS}.
        \item The first is located at $q_z$ = 1.7 \angstrom$^{-1}$,
        corresponding to a real spacing of 3.7 \angstrom~. The reflection is
        attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
        perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
        reflection will be referred to as R-$\pi$.
        \item A weak intensity line is located at exactly half the $q_z$ value of
        R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a
        real space periodic spacing of 7.4 \angstrom~. This reflection has been
        interpreted  as 2\textsubscript{1} helical ordering of aromatic rings
        along the z axis meaning if the positions of the aromatic rings can
        be traced by a helix, then for each turn in the helix, there should be
        two aromatic rings. For this reason it will be referred to as R-helix.
        \item A third major reflection is marked by a low intensity ring located
        at r = 1.4 \angstrom$^-1$. The real space separation
        corresponds to 4.5 \angstrom~ which is characteristic of the average
        spacing between packed alkane chains. This reflection will be called R-alkanes.
        \item Within R-alkanes, are four spots of higher relative intensity which
        will be called R-spots. All are located $\approx 40$ degrees from the $q_z$ axis
        in their respective quadrants. In many liquid crystal systems this can be
        explained by the tilt angle of the alkane chains with respect to the xy plane. % BJC: References
        \item The final feature corresponds to the spacing and symmetry of
        the d\textsubscript{100} plane which can be related to the distance between
        pores. The feature, which will be called R-pores, is characterized by dots
        along $q_z$ = 0. The spacing between dots is indicative of the hexagonal
        symmetry of the packed pores. The same information at higher resolution is obtained using a SAXS
        setup. By radially integrating the 2D data one gets a 1D curve which is 
        shown in Figure~\ref{fig:SAXS}.  % BJC: I also have 2D SAXS data
  \end{itemize}

@@

%MRS2: make it clear these choices are for INITIAL spacing, and that this adjusted based on MD.  Also want to make
sure the dependence of the initial choices is described in supporting information, as the results shouldn't be TOO dependnt on thesechoices.
  We chose the initial layer spacing based on R-$\pi$.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be parallel to the xy plane.
%MRS2: seems like the choice of separation distance here should be detailed in the supporting information, and how it changes  the structures formed.  Also, not quite clear. 
    \item The layers are placed so aromatic rings are stacked 3.7 \angstrom~
    apart in the z-direction.
    \item To quantify the degree of layering in our system, we calculated
    a spatial correlation function, $g(z)$, measured along the z-axis 
    (perpendicular to the membrane plane)
    \item To calculate $g(z)$, we binned the z component distances between 
    the center of mass of each benzene ring and all others of the same pore 
    over 50 ns of equilibrated trajectory and then normalized by the average
    number density.
    \item We compare the degree of layering between systems based on the
    amplitude of the first peak in $g(z)$. We halve the difference between the
    maximum of the first peak and the following minimum. We compare the difference
    to the mean to get a percentage deviation from the average number density.
    \item To extract the average distance between layers we applied a discrete 
    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}
